---
title: "R Notebook"
output:html_document:
  #df_print: paged
  
---
```{r}
# Team Introduction:

# Group No: 10
# Section : 004

# Team Members : 

# 1.Jibina Francis 
# 2.Mankan,Divyajot Singh 
# 3.Srilakshmi Gummadidala


# Background/Motivation :

# Amazon, one of the world's largest online retailers, has had a significant impact on our daily lives. Amazon's vast selection of products and its fast delivery options have made shopping more convenient and efficient for many people. With this dataset, we can gain insights into the sales trends of various products on Amazon over time. By analyzing the data, we can identify which categories of products are the most popular, what factors influence customer ratings and reviews, and how product prices have changed over time, which would help Amazon to improve their products and services.

# Problem Statement :

# Amazon E-commerce platform need a better insight into sales trends of various products sold and customer behavior or market conditions. Specifically, by exploring the relationships between product attributes and their impact on sales, as well as identifying top-performing products and their characteristics. This will Identify factors that contribute to the success of these products by optimizing its sales strategy and drive overall revenue growth.

# Problem Proposal :

# Our project team will create a predictive analytics model to examine the Amazon sales trends on various products sold on their platform to gain knowledge on their overall forecast sales performance. Our team will aim to provide actionable insights that could help Amazon drive revenue growth and offer a better shopping experience to its customers by optimizing their sales strategy.


# Analysis Questions :

# Which Category capture the market percent ?
# Number of Products having high demand in each category ?
# Demand relation with discount ?
# What is the relationship between product ratings and sales of the product in each category?
# Which product categories have the highest and lowest average prices?


# Dataset Description :

# The actual Amazon dataset is having the data of 1K+ Amazon product in different categories along with ratings and reviews as per their details listed on the official website of Amazon.com.
# We have used subset of the actual amazon sales dataset with total six categories in it and sub classified the categories column variables.Total 573 Rows and 16 Columns in the currently used subset.

#VARIABLES:

# PRODUCT_ID : PRODUCT ID
# PRODUCT_NAME : NAME OF THE PRODUCT
# CATEGORIES : CATEGORY OF THE PRODUCT
# DISCOUNTED_PRICE : DISCOUNTED PRICE OF THE PRODUCT
# ACTUAL_PRICE : ACTUAL PRICE OF THE PRODUCT
# DISCOUNT_PERCENTAGE : PERCENTAGE OF DISCOUNT FOR THE PRODUCT
# RATING : RATING OF THE PRODUCT
# RATING_COUNT : NUMBER OF PEOPLE WHO VOTED FOR THE AMAZON RATING
# ABOUT_PRODUCT : DESCRIPTION ABOUT THE PRODUCT
# USER_ID : ID OF THE USER WHO WROTE REVIEW FOR THE PRODUCT
# USER_NAME : NAME OF THE USER WHO WROTE REVIEW FOR THE PRODUCT
# REVIEW_ID : ID OF THE USER REVIEW
# REVIEW_TITLE : SHORT REVIEW
# REVIEW_CONTENT : LONG REVIEW
# IMG_LINK : IMAGE LINK OF THE PRODUCT
# PRODUCT_LINK : OFFICIAL WEBSITE LINK OF THE PRODUCT

```


## Amazon Sales Dataset : Data Visualization using ggplot2


Academic Integrity - Integrity means to make choices according to a certain set of values. Academic integrity means to be guided by shared values and norms of behavior in all aspects of one’s academic work. St. Clair College is committed to the six fundamental values of academic integrity: Honesty, Trust, Fairness, Respect, Responsibility, and Courage. Applying these values can help guide students to make good choices about how to approach their schoolwork. 

Academic integrity means putting those values into practice and being accountable. Faculty, staff and students each play an important role in building and maintaining a culture of academic integrity. All members of our college community have a shared responsibility to uphold the principles of academic integrity so that the teaching-learning relationship is built on mutual trust and respect, and that we maintain an honest and fair learning environment. Academic integrity is integral to a student’s academic success. The habits that students develop on their academic journey while at St. Clair College will carry them not only through their program of study but also into the workplace.


#### R Version :
##### R version 4.2.2 (2022-10-31 ucrt)



#### R Studio Version :  
##### RStudio 2022.12.0+353 



#### Package Versions : 

##### 1.tidyverse - 1.3.2
##### 2.ggplot2 - 3.4.1
##### 3.dplyr - 1.0.10
##### 4.plotly - 4.10.1
#####  5.ggthemes - 4.2.4




```{r}
library(readxl)
installed.packages("readxl")
Amazon_Subset4 <- read_xlsx("C:/Users/Sreelu/OneDrive - St. Clair College/Documents/St Clair/DAB501/Datasets/Amazon_Subset4.xlsx")
```

```{r}
Amazon_Subset4
```

```{r}
length(unique(Amazon_Subset4$categories))
```


#### DESCRIPTION OF THE DATASET : 

##### Summary information about the dataset:

The actual Amazon dataset is having the data of 1K+ Amazon product in different categories along with ratings and reviews as per their details listed on the official website of Amazon.com.
We have used subset of the actual amazon sales dataset with total six categories in it and sub classified the categories column variables.


#### *CHANGES MADE TO THE ORIGINAL DATASET*

We have used subset of the actual amazon sales dataset with total six categories in it and sub classified the categories column variables.
Total 573 Rows and 16 Columns in the currently used subset.

#### Modified from Category column to Categories column variable:

CATEGORIES : CATEGORY OF THE PRODUCT

Six Categories sub classified from total Categories:

##### 1.InputDevices
##### 2.LaptopAccessories
##### 3.Electronics
##### 4.Cables&Adapters
##### 5.Home&Kitchen
##### 6.OfficeProducts



```{r}
library(tidyverse)
install.packages("ggplot2") 
library(ggplot2)
library(dplyr)
library(scales)
library(ggthemes)
```



#### First plot displaying the distribution of a single continuous variable

Discount percentage of various products based on different categories in Amazon sales dataset. 

#### Description : 

Compare to other categories, OfficeProducts which are having under 25% discount rate and sold high in number and InputDevices which were sold more having 25% to 50% discount rate. Cable&Adapters category has the lowest sold rate as compared to other categories with average discount rate.However both Home&Kitchen and Laptop Accessories have highest discount rate as 50% to 75% but sold average number of products. 

#### Plot: 1.1


```{r}
ggplot(Amazon_Subset4, aes(x=discount_percentage, color = categories, fill = categories)) +
  geom_density(alpha=0.7) +
  labs(title="Density Plot for Discount Percentage", caption="Source: Amazon Sales Dataset",
       x="Discount Percentage \n Figure: 1.1", y="Count") +
  theme_bw() +
  theme(plot.title=element_text(size=16, face="bold", hjust=0.5, margin=margin(10,0,20,0)),
        plot.subtitle=element_text(size=12, margin=margin(0,0,20,0)),
        axis.title=element_text(size=12, face="bold"),
        axis.text.x=element_text(angle=65, vjust=0.8),
        axis.ticks.x=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        panel.border=element_blank(),
        panel.background=element_blank(),
        legend.position="right")

```




#### Second plot displaying the distribution of a single continuous variable

Rating given by customers for various products based on different categories in Amazon sales dataset.

#### Description : 

Majority of the ratings of each product ranges from 3.5 to 5.


#### Plot: 1.2


```{r}
Amazon_Subset4$rating <- as.numeric(Amazon_Subset4$rating)
ggplot(Amazon_Subset4) +
  geom_bar(aes(x=rating), fill = "tomato2", bins = 20, alpha = 0.8) +
  labs(title = "Bar Chart for Rating of Various Products",
       caption = "Source: Amazon Sales Dataset",
       x = "Rating \n Figure: 1.2", y = "Count") +
  theme_bw(base_size = 14) +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 14),
        plot.caption = element_text(hjust = 0),
        axis.text.x = element_text(angle = 65, vjust = 0.8, hjust = 1),
        panel.background = element_rect(fill = "white", color = "grey80"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.spacing = unit(1, "lines")) +
  xlim(0,5)
  scale_x_continuous(labels = function(x) format(x, digits = 1))

```




#### First plot displaying the distribution of a single categorical variable


Count of different sold products based on categories in Amazon sales dataset.

#### Description :

Home and Kitchen category is having the highest number of sales as compared to the other categories. Moreover, Laptop Accessories and Office Products are almost having the same number of sales.


#### Plot: 2.1


```{r}
ggplot(Amazon_Subset4)+
  geom_bar(mapping=aes(x=categories),color="white", fill="red",width = 0.8, stat="count")+
   labs(title="Bar Chart for different categories", 
           caption="Source: Amazon Sales Dataset",
        x = "Categories \n Figure: 2.1", y = "Count") +
      theme(axis.text.x = element_text(angle=65, vjust=5))+
  theme_grey(12)+
  theme(plot.caption= element_text(size=10,
                               color="purple",
                                   face="bold",
                                  vjust = 5))
```


#### Second plot displaying the distribution of a single categorical variable

Actual price ranges of various products of Amazon sales data.

#### Description :

Majority of the products which falls under 5000 range of actual price having more sales than any other ranges.



```{r}
library(magrittr)
```



#### Plot: 2.2(1)

```{r}
Amazon_Subset4 <- Amazon_Subset4 %>%
  mutate(category = cut(actual_price, breaks = c(-Inf, 100, 500, 5000, Inf), labels = c("low", "medium_low", "medium_high", "high")))

ggplot(Amazon_Subset4) +
  geom_bar(mapping = aes(x = category, y = ..count..), color = "white", fill = "purple", bins = 10, alpha = 1) +
  scale_x_discrete(labels = c("Low", "Medium_low", "Medium_high", "High")) +
  labs(title = "Bar Chart of Actual Prices",
       subtitle = "Amazon Sales Dataset",
       x = "Actual Price Category \n Figure: 2.2(1)",
       y = "Count",
       caption = "Source: Amazon Sales Dataset") +
  theme(axis.text.x = element_text(angle = 65, vjust = 0.8),
        plot.title = element_text(family = "Arial", size = 20, face = "bold"),
        plot.subtitle = element_text(family = "Arial", size = 16),
        plot.caption = element_text(family = "Arial", size = 12, face = "italic"),
        plot.background = element_rect(fill = "white"))
```

#### Another visualisation plot displaying the distribution of a single categorical variable

Actual price ranges of various products of Amazon sales data.

#### Description : 

Majority of the products which falls under 5000 range of actual price having more sales than any other ranges.


#### Plot: 2.2(2)

```{r}
g <- ggplot(Amazon_Subset4, aes(categories))
g + geom_density(aes(fill=factor(categories)), alpha=0.8, color="white") + 
    labs(title="Density Plot for Product Categories", 
         subtitle="Distribution of Products by Category",
         caption="Source: Amazon Sales Dataset",
         x="Product Categories \n Figure: 2.2(2)",
         y="Density",
         fill="Product Categories") +
    theme_minimal() +
    theme(axis.text.x=element_text(angle=65, vjust=0.8),
          plot.title=element_text(size=16, face="bold", hjust=0.5),
          plot.subtitle=element_text(size=14, hjust=0.5),
          plot.caption=element_text(hjust=0),
          legend.position="right")

```


#### Plot displaying information about both a continuous variable and a categorical variable

Average discount percentage of products for different categories.

#### Description : 

Cable&Adapters and laptop accessories are having above 50% of the average discount rate compared to other categories such as Electronics, Home&kitchen which is having average of below 50%. However, office products are having the least discounted percentage compared to any other category.



#### Plot: 3



```{r}
g <- ggplot(Amazon_Subset4, aes(categories, discount_percentage))
g + geom_boxplot(varwidth=T, fill="plum", color="black", outlier.color="red", outlier.fill="white") + 
    labs(title="Box plot for discounted percentage of categories",
         caption="Source: Amazon Sales Dataset", x = "Categories \n Figure: 3",
    y="Discount Percentage")
```




#### First plot displaying relationship between two variables

The actual price ranges of each product in different categories.

#### Description :

The products which falls under the actual price range of 2000 in most of all categories are higher compared to the other price ranges.



#### Plot: 4.1

```{r}
ggplot(Amazon_Subset4,aes(x=categories,y=actual_price, color = categories))+
   geom_jitter(size=2, width=0.7, alpha = 1)+ 
  geom_smooth(color="red", method="loess", se=FALSE) +
  coord_flip() +
   labs(title="Jitter plot for Actual Price of Various Categories", 
                  caption="Source: Amazon Sales Dataset", y = "actual_price \n Figure: 4.1", x = "categories", 
              ) +
      theme(axis.text.x = element_text(angle=65, vjust=0.8))+
      scale_y_continuous(limits=c(1000,50000))
```





#### First plot displaying relationship between two variables

The actual price ranges of each product in different categories.

#### Description : 

Almost all products in each category are having the same average rating of 4 or more.



#### Plot: 4.2


```{r}
library(ggthemes)
```


```{r}
g <- ggplot(Amazon_Subset4, aes(categories, rating, color = categories)) +
  geom_jitter(width = 0.2, size = 2.5, alpha = 0.7) +
  labs(title="Jitter Plot for Product Categories and Ratings", 
       caption="Source: Amazon Sales Dataset",
       x="Categories \n Figure: 4.2",
       y="Ratings") +
  theme_few(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "right",
    panel.grid.major = element_line(colour = "gray90"),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    axis.line = element_line(colour = "gray60")
  )
g
```






#### Plot with facet grid which is displaying information about both a continuous variable and a categorical variable

Average discount percentage of products for different categories.


#### Description : 

Cable&Adapters and laptop accessories are having above 50% of the average discount rate compared to other categories such as Electronics, Home&kitchen which is having average of below 50%. However, office products are having the least discounted percentage compared to any other category.


#### Plot: 5


```{r}
ggplot(Amazon_Subset4,aes(x=discount_percentage,y=categories,color=category, fill = rating))+ 
  stat_smooth(method="lm",se=FALSE) +
  geom_point(alpha=0.5) + 
  facet_grid(rows=vars(categories), space = "free",scales ="free") + 
  facet_wrap(vars(categories)) + 
  labs(title="Plot for Discount Percentage of Various Categories", 
           caption="Source: Amazon Sales Dataset", x = "Discount_Percentage \n Figure: 5", y = "Categories") +
      theme(axis.text.x = element_text(angle=65, vjust=5))+
  theme_bw()

```




```{r}
library(plotly)
```



#### Interactive plot using ggplotly to display information about discount percentage of various categories.

#### Description : 

Cable&Adapters and laptop accessories are having above 50% of the average discount rate compared to other categories such as Electronics, Home&kitchen which is having average of below 50%. However, office products are having the least discounted percentage compared to any other category.


#### Plot: 6 (Competition Plot)

```{r}
p <- ggplot(Amazon_Subset4,aes(x=categories,y=discount_percentage,color=categories, text=paste("Category: ", categories, "<br>", "Discount Percentage: ", discount_percentage, "%"))) +
  geom_point()+
   labs(title="Interactive Plot for Discount Percentage of Various Categories", caption="Source: Amazon Sales Dataset", x= "Categories \n Figure: 6") +
      theme(axis.text.x = element_text(angle=65, vjust=0.8),
            plot.title = element_text(size = 16),
            plot.caption = element_text(size = 10))
ggplotly(p, style = list(legend.position = "bottom"),
          tooltip = c("x", "y", "text"))

```


##### References to any external sources used and how they were used:

###### Class Lectures/Labs and Data camp courses.

###### Cheat sheet for ggplot: https://www.maths.usyd.edu.au/u/UG/SM/STAT3022/r/current/Misc/data-visualization-2.1.pdf




#### Questions:

##### Q: In what ways do you think data visualization is important to understanding a data set?

###### Data visualization can be extremely useful for understanding any dataset in a number of ways such as 
###### 1.Understanding product performance.
###### 2.Identifying customer behavior.
###### 3.Evaluating marketing campaigns
###### 4.Analyzing customer reviews
###### 5.Monitoring supply chain performance

###### Overall, data visualization provided valuable insights into our Amazon dataset, enabling us to make data-driven decisions that can help improve product performance, customer experience, and supply chain efficiency.




##### Q: In what ways do you think data visualization is important to communicating important aspects of a data set?

###### A: Data visualization is an important tool for communicating important aspects of a data set because it can help to simplify complex data, highlight patterns and trends and to make comparisons which all makes management to have correct decision making.




##### Q: What role does your integrity as an analyst play when creating a data visualization for communicating results to others?

Integrity is important in data visualization because it affects the credibility and trustworthiness of the visualization and the analyst. The audience relies on the analyst to provide an accurate and unbiased representation of the data, and any lack of integrity can cause the audience to lose faith in the visualization and the analyst. This can have serious consequences, particularly if the visualization is being used to make important decisions.




##### Q: How many variables do you think you can successfully represent in a visualization? What happens when you exceed this number?



###### A: The number of variables that can be successfully represented in a visualization depends on the type of visualization and the complexity of the data. In general, it is best to keep the number of variables to a minimum to avoid overwhelming the audience and to ensure that the visualization is easy to read and understand.

###### It is often recommended to limit the number of variables in a visualization to no more than 5 or 6. This is because the more variables that are included in a visualization, the more difficult it becomes to interpret the relationships between them. Additionally, when too many variables are included, the visualization can become cluttered and confusing.






##### Contribution of each team member:



###### There were total 9 plots and each team member have created three plots each out of nine plots.

###### We all 3 team members are responsible for cleaning, processing, and analyzing the data that we used to create the ggplots. We three equally responsible for conducting exploratory data analysis to identify patterns and relationships in the data that can inform the design of the ggplots.

###### We are team members responsible for creating the visual design of the ggplots, including selecting appropriate colors, fonts, and other design elements and also responsible for ensuring that the ggplots are aesthetically pleasing and easy to read.

###### We three equally contributed efforts to prepare the documentation of Problem statement of this Amazon Dataset and other content presented over here in the HTML file.


```{r}
# Filter out extreme outliers
Amazon_Subset4 <- filter(Amazon_Subset4, actual_price <= 20000)

# Filter out additional outliers
Amazon_Subset4 <- filter(Amazon_Subset4, actual_price >= 500)

# Apply a log transformation to actual_price
Amazon_Subset4$log_actual_price <- (log(Amazon_Subset4$actual_price))

# Create a ggplot object
ggplot(Amazon_Subset4, aes(x = log_actual_price)) +
  # Add a histogram layer
  geom_histogram(color = "black", fill = "red", binwidth = 0.5) +
  # Set the x-axis label and title
  xlab("Log Actual Price") +
  ggtitle("Histogram of Log Actual Prices") +
  # Set the theme
  theme_grey(12) +
  # Set the caption
  labs(caption = "Source: Amazon Sales Dataset") +
  # Rotate x-axis labels
  theme(axis.text.x = element_text(angle = 65, vjust = 0.6, hjust = 1))

```
```{r}
sd(Amazon_Subset4$log_actual_price)
```
```{r}
library(ggplot2)

# Calculate the proportion of products in each category
prop_table <- prop.table(table(Amazon_Subset4$categories))
prop <- prop_table

# Create a data frame with the proportion values and category names
# Create a data frame with the proportion values and category names
prop_df <- data.frame(categories = names(prop_table), prop = prop_table)

# Create a pie chart with ggplot2
ggplot(prop_df, aes(x = "", y = prop, fill = categories)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  scale_fill_manual(values = rainbow(length(unique(prop_df$categories)))) +
  labs(title = "Proportions of Products in Different Categories",
       x = NULL, y = NULL, fill = "Categories",
       caption = "Source: Amazon Sales Dataset") +
  theme_void()

```

```{r}
Amazon_Subset4$categories
```
```{r}
#install.packages("prettydoc")
#lightweight template
```



```{r}
ggplot(Amazon_Subset4)+
  geom_bar(mapping=aes(x=discount_percentage),color="white", fill="red",width = 0.8, stat="count")+
   labs(title="Bar Chart for different categories", 
           caption="Source: Amazon Sales Dataset",
        x = "Categories \n Figure: 2.1", y = "Count") +
      theme(axis.text.x = element_text(angle=65, vjust=5))+
  theme_grey(12)+
  theme(plot.caption= element_text(size=10,
                               color="purple",
                                   face="bold",
                                  vjust = 5))


```

